
\newlist{statements}{enumerate}{1}
\newlist{equations}{enumerate}{1}
\setlist[statements,1]{label=\arabic*., left=0.5cm}
\chapter{Mathematical models}
Because our dataset was limited, we used linear regression to try to derive a function that explains how a possible connection 
between user performance and our independent variables looks. We decided against machine learning because it requires much more
data to be effective. More observation per independent variable reduces the chances of overfitting and increases the reliability of a regression
model. There even exist evidence-based guidelines that prescribe having at least 10 observations per independent variable \citep*{Harrell.2015}.
Considering this, we chose the four described variables with their different instances because they ensured having enough observations per
variable.

\section{Aggregated Variables}
Because our study used two nominal scaled variables and we want to analyze the results using linear regression we follow the process
of building aggregated variables. The concept of building composite indicators or variables when facing categorial data is widely used
\citep*{MichelaNardo.2005, Lauro.2018}. Those aggregated variables encode the different instances of the nominally scaled variables into
continuous variables, which in turn enables linear regression. This requires existing knowledge, preferably empirically proven.
\section{Answer time}
First, we want to look at the user performance in terms of their answer time. Because we have so little data available and
because it simplifies the analysis process, we decided to include all data points, even if the answer was incorrect.
Inspired by "Fitts‘ Law", which is often used in the research field on human-computer interaction (HCI) we focus on one possible aspect of answer
time. MacKenzie enhanced the core considerations of Fitts declaring the movement time to be dependent on the difficulty of the task \citep*{MacKenzie.1992}.
We transfer this consideration to our answer time variable and propose an index that describes the level of difficulty of the task to which
the user has answered. Besides the difficulty index, there are $n-1$ more factors that may influence answer time.
\begin{equation} \label{answerTimeEquation}
    Answer_{time} = a_0 + a_1 * Index_{difficulty} + a_2 * X_2 + \dots + a_n * X_n
\end{equation}

\begin{conditions}
    Index_{difficulty}     & difficulty of a specific question using one specific dashboard variant  \\
    a_0                    & intercept \\
    a_1                    & coefficient of the difficulty index \\
    X_2\dots X_n           & other factors that also influence the answer time \\
    a_2\dots a_n           & coefficients of other factors \\
\end{conditions}
We now want to look how all our independent variables influenced the difficulty index.
\subsection{Number of views}
One of the rationales of MCV is that more views on the same data, help with data exploration. As we have already seen if MCV
is applied correctly, cognitive overhead is reduced compared to the same application not using multiple views.
Because comparison is a type of data exploration we can assume that more views also help increase user performance in tasks
that involve comparison. We can assume:
\begin{statements}
    \item A higher number of views on the same data will reduce the difficulty index.
\end{statements}
\subsection{Quality of views}
The suitability of views changes across different contexts. One context for example is what task is tried to be solved
using the view. In our experiment, we used a total of four different task types, which are represented by the four question
types. Lohse et al. classify visual representations(views)
depending on the type of information conveyed. They classified 11 different view types and empirically derived Likert
scores (1-9) for every type of information \citep*{Lohse.1994}. Among other things, they classified \textit{time charts} and 
\textit{tables} when dealing with \textit{temporal} or \textit{numerical} information. As the \textit{attribute in space} operand
asks for \textit{numerical} information and the \textit{space in time} operand asks for \textit{temporal} information, we can 
use these empirical Likert scores to quantify the quality of the table and time chart used when answering a specific question.
Since each interaction technique renders at least one table and one time chart and we cannot predict which view the user will use, we
need to calculate the mean view quality of both views for a given question type. The Likert scale ranged from one to nine, meaning if a
visualization is rated as a nine, it conveys the information one hundred percent. In our calculation, we divide the Likert scores by nine to account
for this. The following equation uses the Likert scores from the tables and the time chart visualizations:
\begin{equation} \label{qualityViewEquation}
    q_{view}= 
    \begin{cases}
        \frac{8.0 / 9 + 4.5 / 9}{2} \approx 0.69,& \text{if } question\_type == attribute\_in\_space \\
        \frac{1.8 / 9 + 7.8 / 9}{2} \approx 0.53,& \text{if } question\_type == space\_in\_time 
    \end{cases}
\end{equation}
Questions could also be of type \textit{identify} or \textit{measure}.
In our mathematical model, we do not make any assumptions about \textit{identify} or \text{measure} questions because we lack research that
would help quantify view quality in those regards. We can conclude with the following assumption:
\begin{statements}[resume]
    \item A higher encoding quality of the views regarding the type of question will reduce the difficulty index
\end{statements}
\subsection{Distraction index}
Finally, we look at the concept of distraction. Through personal observation, we concluded that the key difference between our three interaction
techniques can be found in the prevailing distraction. In all our questions we want to find specific values of our spatio-temporal data. Therefore
all other possible values on the dashboard interface can act as \textit{distractors}. Distractors have ranging distraction impact depending on
their \textit{similarity} to the searched value. When a user starts the comparison process the effect of the different interaction techniques
takes place. We can count the number of distractors displayed that could theoretically be confused with the value that is looked for. We can
multiply the distractors with a weight that represents the impact of each distractor. Because all interaction techniques follow the principle
of visually emphasizing the selected target states from the remaining states we have decided to distinguish between two cases. From now on we classify
each distractor as either \textit{similar distractor} containing all distractors that look similar to the target value or as
\textit{background distractor} containing all distractors that look different from the target value but still exist on the dashboard
interface and could also be mistaken to be the target value. As we lack research for quantifying the weights of the different distractors, we
decided that the similar distractors are twice as distracting as the background distractors. 

To compute the overall distraction, all similar distractors
and background distractors are counted and multiplied by their respective weight resulting in a value that has to be divided by two because a user
will only use one of the two views (table view \&\ time chart view) simultaneously and both views encode the same information. This is the same principle
described earlier for the view quality. The resulting value is called the \textit{distraction index}. The number of comparison targets also impacts
the distraction because now more values are similar distractors, resulting in a higher distraction index. This also supports our earlier presented finding
of a higher number of comparison targets increasing the difficulty of the comparison process. The distraction index is further influenced by the question
type, our third independent variable. Because \textit{measure} and \textit{space in time} both ask for multiple target values that are required to answer the
question, the user has to search multiple times. In this case, everything is counted the same only with reduced \textit{similar distractors}. We also account
for the multiple searches the user has to perform by adding the not-yet-found target values in every search to the distraction index, getting less with every
completed search.
Finally, the number of views in connection with the question type also influences the distraction index. Summarizing we can say that the additional
views on the right side of the dashboard variants with the four views add similar distractors and background distractors depending on the question type.
Questions with the question type \textit{measure} ask for differences between comparison targets and therefore the additional views on the right side also contain
target values along with some \textit{similar distractors}. On the other hand, questions that are of type \textit{identify} never ask for differences between
comparison targets, which is why the additional views only add background distractors. We can compute the distraction each view generates using this formula:
\begin{equation} \label{viewDistractionEquation}
    View_{distraction} = \alpha * (N_{distractor\_similar} * \beta + N_{distracor\_background} * \gamma)
\end{equation}
\begin{conditions}
    N_{distractor\_similar}     &  number of similar distractors \\
    N_{distractor\_background}  &  number of background distractors \\   
    \alpha                      &  coefficient of the distraction of a view \\
    \beta                       &  multiplier for the similar distractors \\
    \gamma                      &  multiplier for the background distractors \\
\end{conditions}
In the analysis, we will set $\beta = 1$ and therefore $\gamma = 0.5$. The coefficient $\alpha$ will be manipulated in the analysis later. 
We calculate the distraction each of the four possible views generates. We abbreviate the four views by their absolute position on the dashboard
(UL = upper left, LL = lower left, UR = upper right, LR = lower right). Like on the view quality, we need to calculate the mean of the distraction scores
of the table view and the time chart view, because one user will search and find the information in only one of the two views. The final formula for
calculating the distraction index looks like this:
\begin{equation} \label{distractionIndexEquation}
    Index_{distraction} = \frac{UL_{distraction} + LL_{distraction} + UR_{distraction} + LR_{distraction}}{2}
\end{equation} 
Because the distraction index encapsulates the total distraction the user experiences when answering a comparative question on one of the dashboard variants,
we can derive:
\begin{statements}[resume]
    \item A higher distraction index will increase the difficulty index.
\end{statements}

\subsection{Difficulty index}

The aggregation of all of our variables and the three numbered assumptions we derived lead to following formula discribing the difficulty index:

\begin{equation} \label{difficultyIndexEquation}
    Index_{difficulty} = b_0 + \frac{b_1}{n_{view}} + \frac{b_2}{q_{view}} + b_3 * Index_{distraction}
\end{equation}


\section{Answer accuracy}
As already explained the answer accuracy was measured in binary and therefore multiple linear regression is not an option. Instead, we will use the
method of logistic regression to analyze the results. When thinking of factors that could influence the accuracy of an answer we made the
following assumptions. Because answer time does not matter for the accuracy we concluded that no factor can impede the accuracy of the answer besides the
distraction that is present on the dashboard. A user can mistakenly confuse the correct answer with all distractors present. We call this variable 
discriminability. As with the answer time, there are more unknown and unobserved variables that have a varying impact on the answer accuracy.
We conclude:

\begin{equation} \label{answerAccuracyEquation}
    Answer_{accuracy} = a_0 + a_1 * Index_{discriminability} + a_2 * X_2 + \dots + a_n * X_n
\end{equation}

\begin{conditions}
    Index_{discriminability}  & discriminability of the target value \\
    a_0                       & intercept \\
    a_1                       & coefficient of the discriminability index \\
    X_2\dots X_n              & other factors that also influence the answer accuracy \\
    a_2\dots a_n              & coefficients of other factors \\
\end{conditions}

\subsection{Discriminability index}
The discriminability is the multiplicative inverse of the distraction index. When no distractor is present
the user will always answer every question correctly. With a rising number of distractors the chance of mixing up the target values is increasing.
Like before we have to account for different distraction qualities. Different distractors have different distracting qualities because of their visual appearance.
It is the same observation that is covered with the distraction index. Because of this, we will reuse it in the formula for the discriminability index:
\begin{equation} \label{discriminabilityIndexEquation}
    Index_{discriminability} = \frac{1}{1 + Index_{distraction}}
\end{equation}
